---
title: "Part 1"
author: Malley Smith 
output: 
  html_document:
    code_folding: hide
    df_print: paged
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(RSQLite)
library(DBI)


```

Template file. Code will be included in folded blocks in the output to facilitate grading. Please knit this file and commit both the rmd and the html output. If you add external files to your analysis, please commit them to the files folder in this repository. NOTE: please do not commit large (15MB+) files to GitHub. Instead please denote the origin of the files in your code. 

```{r}
#example connection to database. note that you have to download the database from onedrive

con <- DBI::dbConnect(RSQLite::SQLite(), "/Users/apple/Library/CloudStorage/OneDrive-UniversityofIllinoisatChicago/GitHub/database/detroit.sqlite")


# sales tbl

dplyr::tbl(con, 'sales')

# convert to tibble
dplyr::tbl(con, 'sales') %>% dplyr::collect()

# sql query

dplyr::tbl(con, 'sales') %>% count(year(sale_date))

# dplyr::tbl(con, 'sales') %>% count(year(sale_date)) %>% show_query()

```

Section A: Exploratory Data Analysis



```{r}
# Getting familiar with the database 
# src_dbi(con) #tbls: assessments, blight, foreclosures, parcels, parcels_historic, sales


# Creating individual tables 
###########################
assessments <- tbl(con, "assessments")
head(assessments, n = 20)

blight <- tbl(con, "blight")
head(blight, n = 10)

foreclosures <- tbl(con, "foreclosures")
head(foreclosures, n = 10)


parcels <- tbl(con, "parcels")
head(parcels)

parcels_historic <- tbl(con, "parcels_historic")

sales <- tbl(con, "sales")
head(sales)

##########################

dplyr::tbl(con, 'sales') %>% count(year(sale_date)) %>% arrange(desc(n)) 
# 2014 has largest number os sales 


sales %>% 
  group_by(year(sale_date)) %>%
  summarise(mean_sale_price = mean(sale_price, na.rm = TRUE), median = median(sale_price, na.rm = TRUE)) %>%
  arrange(desc(mean_sale_price))

# 2020 has the highest average sale price


library(lubridate)  


# Joining assessment and sales tables, joinging by parcel number 
sales_joined <- sales %>%
  left_join(assessments, by=c('parcel_num' = 'PARCELNO'))
  

```


Data Quality 
# Foreclosures is a wide data set where NA means the property is not foreclosed and 1 = foreclosed. 


```{r}
# Overview

# The thought here is that the assessed value is more than what the property sells for. So the sale price is subtracted from the assessed value 

sales_joined2 <- sales_joined %>% 
  mutate(difference = ASSESSEDVALUE - sale_price)

sales_joined2 %>% 
  group_by(year) %>%
  summarise(mean_difference = mean(difference, na.rm = TRUE), median_difference = median(difference, na.rm = TRUE), min = min(difference, na.rm = TRUE), max = max(difference, na.rm = TRUE)) %>%
  arrange(desc(difference))

  
# So far 2022 has the largest positive mean difference, meaning that properties on average are selling for less than their assessed value. Because there is less than two months worth of data for 2022, it will not be a helpful comparison. This may not be a good year to keep if we want to compare full years. 2011 is the second largest positive mean difference. 


# 2017 has the largest negative mean difference. Meaning on average properties were sold above their assessed value. 


```

When comparing this brief analysis of Detroit's Property Assessment Data, the first impression of the data when only looking at the means is that properties generally are being assessed at a lower value than their sale price. 

However, if we look at the median, minimum and maximum. We see that medians are all positive. It seems like there are still a good amount of properties that have larger assessed values compared to their sale price. What this means is that t

Section B: Sales Ratio Study 

```{r}
devtools::install_github("cmf-uchicago/cmfproperty")
library(cmfproperty)

sales_joined3 <- sales_joined %>% 
  select(parcel_num, year, sale_price, ASSESSEDVALUE)

# df <- cmfproperty::sales_joined3

# Not sure how exactly I am supposed to use this package. I thought the line above would create my object in cmfproperty but instead it does not recognize the object sales_joined3. 

sales_ratio <- sales_joined3 %>%
  mutate(sales_ratio = ASSESSEDVALUE/sale_price)

sales_ratio %>% 
  group_by(year) %>%
  summarise(mean_sr = mean(sales_ratio), median_sr = median(sales_ratio))

```

Section C: Property Sales Simple Regression 

```{r}
 
sales_mod1 <- lm(sale_price ~ ASSESSEDVALUE, data = sales_joined)
summary(sales_mod1)
# plot(sales_mod1)


sales_mod2 <- lm(sale_price ~ ASSESSEDVALUE*year, data = sales_joined)
summary(sales_mod2)



sales_mod3 <- lm(ASSESSEDVALUE ~ year + sale_price, data = sales_joined)
summary(sales_mod3)
  


```



Section D: Foreclosures Simple Regression 

```{r}

foreclosures2 <- foreclosures %>%
  pivot_longer(!c(prop_addr, prop_parcelnum),
             names_to = "year",
             values_to = "foreclosure")

foreclosures2 %>%
  group_by(year) %>%
  summarise(n_foreclosures = count(foreclosure)) %>%
  arrange(desc(n_foreclosures))
# 2015 has the most foreclosures




forclosure_mod1 <- lm(foreclosure ~ year, data = foreclosures2)

```








